Predicting Internet Path Dynamics and Performance with Machine Learning

### 2017-12-20 paper-tex/paper.tex
resume

In the original paper, the authors focus on the problem of predicting Internet path changes and path performance using \texttt{traceroute} measurements. They use the recent route information of paths (route age for paths, route changes in the past, average RTT, etc) to do some predictions of paths. Their are three predict targets: $(i)$ the remaining life time of a route (i.e., the time before a route changes), $(ii)$ the number of route changes in a future timeslot, and $(iii)$ the average RTT  that the path $P$ will experience in the next \texttt{traceroute} measurement.

@@

We find some drawbacks in their data process.(see section \ref{sec:dataAna} )And in our own work,firstly we reprocess the provided data in more resonable ways and use the same method (random forest) mentioned by the author to compare the results. Secondly, two other machine learning methods , xgboost and LSTM, are applied to get better performance.

\section{Related Works}

1. introduce the original paper

In the original paper, the problem they solve is to use random forest model to predict three labels for one path: $(i)$ the remaining life time of a route for one path at time $t$ (i.e., the time before a route changes), namely $R_r(t)$, $(ii)$ the number of route changes for one path $P$ in a future timeslot $T$, namely $rcP_T$  and $(iii)$ the average RTT that the path $P$ will experience in the next \texttt{traceroute} measurement, namely $avgRTT_P(t)$.

@@

And the authors use three different ways of selecting input features: $(i)$  use all of the 69 features to predict each target, $(ii)$ when predicting target $X$, use corresponding feature set $F_X$, $(iii)$ after wrapper-based feature selection, use the top features for each target. And finally, what they get is the third way of choosing input features can achieve better results (though there is only minor differences in output results).

@@

\section{Data Analysis}
 \label{sec:dataAna}
\hui{I think this is an important part, who will take this part?}
\heng{Plots of temporally differentiated prediction targets are given here to show their behavior patterns in temporal domain, thus supporting the use of LSTMs.}
With careful review of the data used in the original paper, here we (a) show the authors had used the data incorrectly in the context of making \textbf{predictions} and (b) reveal the temporal properties that the original method failed to employ but perfectly fits LSTMs (long short term memory neural networks), which we later use to improve the performance.

\subsection{Justification of Data Processing}
The original paper \cite{predict} made a major mistake on data employment: they implicitly used data from the future to predict events in the future, which is not possible. More specifically, all of their tests on the performance of \textit{NetPerfTrace} depend on statistics calculated out of data of all time, including the data they were to predict. Features like \{5, 15, 25...\} percentage point of route life, for example, takes the route life of all routes during the whole week into account, including route $r_i$. However, these features still take part in predicting the residual life of $r_i$. This is just like guessing something when you knew it.

In fact, a principle of making predictions is to use the data \textbf{already observed} to guess the unobserved. Thus, in all the experiments of this paper, we made it a hard constraint that all the features used to predict any event at, say, time $t$, only come from the raw data collected \textbf{before} $t$. First, we re-validated the performance of classic machine learning methods on the 3 tasks according to our justified data processing manner, in the following 3 ways: (a) use the global features (max value, min value, distribution, etc.) of the former k\% data to train the models and predict the targets of the latter (1-k)\% data, (b) use the global features to train the models and update them during the predicting procedure and (c) use data in a certain recent time-slot to calculate the global features. Then we train LSTMs on these data in two different ways: (a) input only a scalar, the prediction target, to the network and (b) input the prediction target and the global features as a vector.

\subsection{Temporal Properties of the Data}
We believe temporally adjacent routes and \textbf{traceroute} records of a path share some common network nodes, thus their behavior contain implicit dependencies in temporal domain. While the data are organized in the order of time, the original method ignored this property but regard them as independent instances, i.e., the prediction for the next time step only depends on the features of the present, but does not use information given by more previous time steps. This manner fails the original method to capture even the most naive temporal dependencies. Here we do not provide strict proofs of the temporal dependencies that lie in these routes but simply throw some light on the possibility of their existence.

@@

\section{Methodology}
\subsection{Data Processing}
We reprocess the data in several more rational ways, which in other words, predicting targets using the previous information only. And to be specific, we divide the data into two sets, namely training set and test set. We calculate statistical properties (global features) of paths using three different methods: $(i)$ use the global features (max, min, average, percentiles, etc.) in the training set to get the model and predict the targets of the latter $(1-k)\%$ data, $(ii)$ use the global feature of the former $k\%$ to train the models and update these features during the predicting procedure(i.e.,update the global features everytime there is a new \texttt{traceroute} measurement of a path), and $(iii)$ only use recent \texttt{traceroute} data of certain length (namely $window\ size$) in the training set to calculate the global features, and when using the trained model to predict targets, we only consider features calculated in  recent \texttt{traceroute} measurements of the same $window\ size$ . For convenience, in the following article, we use $data-i$ to represent reprocessing the original data using the $i$th method ($i$ can be either 1,2 or 3). And for the third method, we use $data-3-ws\#$ to indicate the $window\ size$ of $\#$.

@@

\section{Experiment}
\subsection{Random Forest Model}
In our experiment, firstly, we use the standard random forest model (the same as the original paper) with the reprocessed data.  In practice, we set $k = 70, 80$ respectively. And in the third method of processing data, the $window\ size$ is set to $10,20$ and $50$ recent \texttt{traceroute} measurements, respectively. We find that there is little difference when setting $k$ either to 70 or 80. Below are the results when $k = 70$. And we also define an evaluation metric $ABC$, which indicates the  area below the curve ($ABC\in[0,1]$). The closer $ABC$ is to $1$, the better prediction result it has. These figures in \textbf{Fig.\ref{fig:task1_RF}} and \textbf{Fig.\ref{fig:task3_RF}} and  illustrate the relationship between sample percentage and relative prediction error(i.e., $\mid Real - Predict\mid/ Real $). 
\begin{figure*} [!htb] 
	%\vspace{-4pt}  
	\centering  
	\includegraphics[width=0.3\textwidth]{fig4/data-1-task-1-7030-RF.jpg}  
	\includegraphics[width=0.3\textwidth]{fig4/data-2-task-1-7030-RF.jpg} 
\end{figure*} 
\begin{figure*} 	
	\includegraphics[width=0.3\textwidth]{fig4/data-3-task-1-7030-windowsize10-RF.jpg} 
	\includegraphics[width=0.3\textwidth]{fig4/data-3-task-1-7030-windowsize20-RF.jpg}
	\includegraphics[width=0.3\textwidth]{fig4/data-3-task-1-7030-windowsize50-RF.jpg}
	%\vspace{-20pt}  
	\caption{Relative prediction errors for $R_r(t)$ using three different data processing methods, and $window\ size$ for the below three figures are 10,20 and 50.}
	\label{fig:task1_RF}  
	%\vspace{-10pt}  
\end{figure*}

\begin{figure*} [!htb] 
	%\vspace{-4pt}  
	\centering  
	\includegraphics[width=0.3\textwidth]{fig4/data-1-task-3-7030-RF.jpg}  
	\includegraphics[width=0.3\textwidth]{fig4/data-2-task-3-7030-RF.jpg} 
\end{figure*} 
\begin{figure*} 	
	\includegraphics[width=0.3\textwidth]{fig4/data-3-task-3-7030-windowsize10-RF.jpg} 
	\includegraphics[width=0.3\textwidth]{fig4/data-3-task-3-7030-windowsize20-RF.jpg}
	\includegraphics[width=0.3\textwidth]{fig4/data-3-task-3-7030-windowsize50-RF.jpg}
	%\vspace{-20pt}  
	\caption{Relative prediction errors for $avgRTT_P(t)$ using three different data processing methods, and $window\ size$ for the below three figures are 10,20 and 50.}
	\label{fig:task3_RF}
	%\vspace{-10pt}  
\end{figure*}

@@

Table 1 describes the $ABC$ in each figure.
\begin{table}[tbp]
	\centering  
	\begin{tabular}{lcc} 
		\hline
		Target &$R_r(t)$ &$avgRTT_P(t)$ \\ \hline
		$data-1$ & 0.4016    &0.8124   \\        
		$data-2$ &0.3985 &0.8261 \\        
		$data-3-ws10$  &0.7132 &0.8018 \\
		$data-3-ws20$ &0.6314 &0.7910 \\
		$data-3-ws50$ &0.5359  &0.7790 \\ \hline
	\end{tabular}
	\caption{The $ABC$ of curves
		 in \textbf{Fig.\ref{fig:task1_RF}} and \textbf{Fig.\ref{fig:task3_RF}} }
	 \label{tab:result_RF}
\end{table}
From \textbf{Tab.\ref{tab:result_RF}}, it can be inferred that for target $avgRTT_P(t)$: there is minor difference using five different methods to process the data. They all achieve a relatively good result, which also indicates that for average RTT reported from \texttt{traceroute} measurement, it is not sensitive to the temporal locality. However, for the first predicting target, $R_r(t)$, it is very sensitive to the temporal locality. It changes rapidly as time changes. And to achieve better prediction result, using too much previous data will result in noise and decrease the performance of the model. Because from \textbf{Tab.\ref{tab:result_RF}}, we can clearly observe that $data-3-ws10$ achieves larget $ABC$ compared with other processing methods. It considers only the recent data in a relatively short time slot. Indeed, it is reasonable to get such conclusion, for the remaining life time of a route is always closely related to their route age (which can be indicated from the previous \texttt{traceroute} data).
\subsection{Deep Models}

>>>>

	In the original paper\cite{oripaper}, the authors focus on the problem of predicting Internet path changes and path performance using \texttt{traceroute} measurements. They use the recent route information of paths (route age for paths, route changes in the past, average RTT, etc) to do some predictions of paths. Their are three predict targets: $(i)$ the remaining life time of a route (i.e., the time before a route changes), $(ii)$ the number of route changes in a future timeslot, and $(iii)$ the average RTT  that the path $P$ will experience in the next \texttt{traceroute} measurement.

@@

	We find some drawbacks in their data process.(see section \ref{sec:dataAna}) And in our own work, firstly we reprocess the provided data in more resonable ways and use the same method (random forest model) mentioned by the original paper\cite{oripaper} to compare the results. Secondly, another machine learning method, LSTM model, is applied to get better performance.

	\section{Related Works}
	In the original paper\cite{oripaper}, the problem they solve is to use random forest model to predict three targets for one path: $(i)$ the remaining life time of a route for one path at time $t$ (i.e., the time before a route changes), namely $R_r(t)$, $(ii)$ the number of route changes for one path $P$ in a future timeslot $T$, namely $rcP_T(t)$  and $(iii)$ the average RTT that the path $P$ will experience in the next \texttt{traceroute} measurement, namely $avgRTT_P(t)$.

@@

	And the authors use three different ways of selecting input features: $(i)$  use all of the 69 features to predict each target, $(ii)$ when predicting target $X$, use corresponding feature set $F_X$, $(iii)$ after wrapper-based feature selection, use the top features for each target. And finally, what they get is the third way of choosing input features can achieve better results (though there is only minor difference in output results).

@@

	\section{Data Analysis}
	\label{sec:dataAna}
	\hui{I think this is an important part, who will take this part?}
	\heng{Plots of temporally differentiated prediction targets are given here to show their behavior patterns in temporal domain, thus supporting the use of LSTMs.}
	With careful review of the data used in the original paper, here we: $(i)$ show the authors had used the data incorrectly in the context of making predictions, and $(ii)$ reveal the temporal properties that the original method failed to employ but perfectly fits LSTM model (long short term memory neural networks), which we will later use to improve the performance.
	\subsection{Justification of Data Processing}
	The original paper \cite{oripaper} made a major mistake on data employment: they implicitly used data from the future to predict events in the future, which is not possible. More specifically, all of their tests on the performance of \textit{NetPerfTrace} depend on statistics calculated out of data of all time, including the data they were to predict. Features like percentiles of route life, for example, take the route life of all routes during the whole week into account, including route $r_i$. However, these features still take part in predicting the residual life of $r_i$. It resembles guessing something you already knew.

	In fact, a principle of making predictions is to use the data already observed to guess the unobserved. Thus, in all the experiments of this paper, we made it a hard constraint that all the features used to predict any event at, say, time $t$, only come from the raw data collected before $t$. First, we re-validated the performance of classic machine learning method on two tasks(task 1 and task 3) according to our justified data processing manner, (The reason why we miss the task 2 is that it is not possible to reprocess data with respect to task 2 according to the raw data given by the original paper) in the following three ways: $(i)$ use the global statistical features (average, minimum, maximum, and percentiles, etc) of the former $k\%$ data to train the models and predict the targets of the latter $(1-k)\%$ data, $(ii)$ use the global features to train the models and update them during the predicting procedure and $(iii)$ use data in a certain recent time-slot to calculate the global features. Then we train LSTMs on these data in two different ways: $(i)$ input only a scalar, the prediction target, to the network and $(ii)$ input the prediction target and the global features as a vector.

	\subsection{Temporal Properties of the Data}
	We believe temporally adjacent routes and \texttt{traceroute} records of a path share some common network nodes, thus their behavior contain implicit dependencies in temporal domain. While the data are organized in the order of time, the original method ignored this property but regard them as independent instances, i.e., the prediction for the next time step only depends on the features of the present, but does not use information given by more previous time steps. This manner fails the original method to capture even the most naive temporal dependencies. Here we do not provide strict proofs of the temporal dependencies that lie in these routes but simply throw some light on the possibility of their existence.

@@

	\section{Methodology}
	\subsection{Data Processing}
	We reprocess the data in several more rational ways, which in other words, predicting targets using the previous information only. And to be specific, we divide the data into two sets, namely training set and test set. We calculate statistical properties (global features) of paths using three different manners: $(i)$ use the global features (max, min, average, percentiles, etc.) in the training set to get the model and predict the targets of the latter $(1-k)\%$ data, $(ii)$ use the global feature of the former $k\%$ to train the models and update these features during the predicting procedure(i.e.,update the global features everytime there is a new \texttt{traceroute} measurement of a path), and $(iii)$ only use recent \texttt{traceroute} data of certain length (namely $window\ size$) in the training set to calculate the global features, and when using the trained model to predict targets, we only consider features calculated in  recent \texttt{traceroute} measurements of the same $window\ size$ . For convenience, in the following article, we use $data-i$ to represent reprocessing the original data using the $i$th manner ($i$ can be either 1,2 or 3). And for the third manner, we use $data-3-ws\#$ to indicate the $window\ size$ of $\#$.

@@

	\section{Experiment}
	\subsection{Random Forest Model}
	In our experiment, firstly, we use the standard random forest model (the same as the original paper) with the reprocessed data.  In practice, we set $k = 70, 80$ respectively. And in the third manner of processing data, the $window\ size$ is set to $10,20$ and $50$ recent \texttt{traceroute} measurements, respectively. We find that there is little difference when setting $k$ either to 70 or 80. Below are the results when $k = 70$. And we also define an evaluation metric $ABC$, which indicates the  area below the curve ($ABC\in[0,1]$). The closer $ABC$ is to $1$, the better prediction result it has. These figures in \textbf{Fig.\ref{fig:task1_RF}} and \textbf{Fig.\ref{fig:task3_RF}} and  illustrate the relationship between sample percentage and relative prediction error(i.e., $\mid Real - Predict\mid/ Real $). 
	\begin{figure*} [!htb] 
		%\vspace{-4pt}  
		\centering  
		\includegraphics[width=0.3\textwidth]{fig4/data-1-task-1-7030-RF.pdf}  
		\includegraphics[width=0.3\textwidth]{fig4/data-2-task-1-7030-RF.pdf} 
	\end{figure*} 
	\begin{figure*} 	
		\includegraphics[width=0.3\textwidth]{fig4/data-3-task-1-7030-windowsize10-RF.pdf} 
		\includegraphics[width=0.3\textwidth]{fig4/data-3-task-1-7030-windowsize20-RF.pdf}
		\includegraphics[width=0.3\textwidth]{fig4/data-3-task-1-7030-windowsize50-RF.pdf}
		%\vspace{-20pt}  
		\caption{Relative prediction errors for $R_r(t)$ using three different data processing manners, and $window\ size$ for the below three figures are 10,20 and 50.}
		\label{fig:task1_RF}  
		%\vspace{-10pt}  
	\end{figure*}

	\begin{figure*} [!htb] 
		%\vspace{-4pt}  
		\centering  
		\includegraphics[width=0.3\textwidth]{fig4/data-1-task-3-7030-RF.pdf}  
		\includegraphics[width=0.3\textwidth]{fig4/data-2-task-3-7030-RF.pdf} 
	\end{figure*} 
	\begin{figure*} 	
		\includegraphics[width=0.3\textwidth]{fig4/data-3-task-3-7030-windowsize10-RF.pdf} 
		\includegraphics[width=0.3\textwidth]{fig4/data-3-task-3-7030-windowsize20-RF.pdf}
		\includegraphics[width=0.3\textwidth]{fig4/data-3-task-3-7030-windowsize50-RF.pdf}
		%\vspace{-20pt}  
		\caption{Relative prediction errors for $avgRTT_P(t)$ using three different data processing manners, and $window\ size$ for the below three figures are 10,20 and 50.}
		\label{fig:task3_RF}
		%\vspace{-10pt}  
	\end{figure*}

@@

	Table 1 describes the $ABC$ in each figure.
	\begin{table}[tbp]
		\centering  
		\begin{tabular}{lcc} 
			\hline
			Target &$R_r(t)$ &$avgRTT_P(t)$ \\ \hline
			$data-1$ & 0.4016    &0.8124   \\        
			$data-2$ &0.3985 &0.8261 \\        
			$data-3-ws10$  &0.7132 &0.8018 \\
			$data-3-ws20$ &0.6314 &0.7910 \\
			$data-3-ws50$ &0.5359  &0.7790 \\ \hline
		\end{tabular}
		\caption{The $ABC$ of curves
			in \textbf{Fig.\ref{fig:task1_RF}} and \textbf{Fig.\ref{fig:task3_RF}} }
		\label{tab:result_RF}
	\end{table}
	From \textbf{Tab.\ref{tab:result_RF}}, it can be inferred that for target $avgRTT_P(t)$: there is minor difference using five different manners to process the data. They all achieve a relatively good result, which also indicates that for average RTT reported from \texttt{traceroute} measurement, it is not sensitive to the temporal locality. However, for the first predicting target, $R_r(t)$, it is very sensitive to the temporal locality. It changes rapidly as time changes. And to achieve better prediction result, using too much previous data will result in noise and decrease the performance of the model. Because from \textbf{Tab.\ref{tab:result_RF}}, we can clearly observe that $data-3-ws10$ achieves larget $ABC$ compared with other processing manners. It considers only the recent data in a relatively short time slot. Indeed, it is reasonable to get such conclusion, for the remaining life time of a route is always closely related to their route age (which can be indicated from the previous \texttt{traceroute} data).
	\subsection{Deep Models}
